High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-07-21 inst/manuscript/manuscript.Rmd
Tweaked abstract, incremented package

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: no
spacing: doublespacing
footerdate: no
note: This pre-print not yet undergone peer review. It has been submitted to the \emph{Frontiers in Artificial Intelligence}. This version will be updated as it is revised, and the final published version will be accessible through its DOI link.   
abstract: Mapping the changing characteristics of Africa's smallholder-dominated agricultural systems, including the sizes and numbers of fields, is key to understanding critical questions related to food security and other pressing socioeconomic and environmental concerns. However, accurately mapping these systems is difficult because of two key problems: 1) the spatial and temporal mismatch between most publicly satellite sensors and smallholder fields; 2) the difficulty of collecting high quality labels for training and assessing machine learning classifiers. We developed a cropland mapping approach designed to address these two problems, which we used to map Ghana’s croplands (excluding those under tree crops) for the year 2018.  To overcome the first problem, we converted daily, high resolution CubeSat (PlanetScope) imagery into two cloud-free seasonal composites covering the 2018 agricultural calendar. To address the second problem, we created a labelling platform that rigorously assesses and minimizes label error, and used it to iteratively train a machine learning classifier (Random Forests) using an active learning approach, in which we used the uncertainty in model predictions to select more informative training samples. Minimizing label errors improved model F1 scores by up to 25%, while active learning increased the F1 score by an average of 9.1% between first and last training iterations, which was 2.3% higher than models trained with randomly selected labels. We used the resulting <4 m map of cropland probabilities within a final segmentation step to delineate crop field boundaries. Using an independent map reference sample (n=1,207), we found that overall accuracies of the cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%. Relying on the map reference sample to generate an unbiased area estimate from the field boundary map, we calculate that cropland covers 17.1 (15.4-18.9%). Using the labellers’ digitized field boundaries to estimate biases in the size and number of segments in the field boundary map, we estimate that there are 1,662,281 crop fields in Ghana with an average size of 1.73 ha. Our results demonstrate an adaptable and transferrable approach for mapping the characteristics of croplands on an annual basis and over national extents, with several features that effectively minimize the errors inherent in remote sensing of smallholder-dominated agriculture.

>>>>

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: no
spacing: doublespacing
footerdate: no
note: This pre-print not yet undergone peer review. It has been submitted to the \emph{Frontiers in Artificial Intelligence}. This version will be updated as it is revised, and the final published version will be accessible through its DOI link.   
abstract: Mapping the changing characteristics of Africa's smallholder-dominated agricultural systems, including the sizes and numbers of fields, is key to understanding critical questions related to food security and other pressing socioeconomic and environmental concerns. However, accurately mapping these systems is difficult because of two key problems: 1) the spatial and temporal mismatch between most publicly satellite sensors and smallholder fields; 2) the difficulty of collecting high quality labels for training and assessing machine learning classifiers. We developed a cropland mapping approach designed to address these two problems, which we used to map Ghana’s croplands (excluding those under tree crops) for the year 2018.  To overcome the first problem, we converted daily, high resolution CubeSat (PlanetScope) imagery into two cloud-free seasonal composites covering the 2018 agricultural calendar. To address the second problem, we created a labelling platform that rigorously assesses and minimizes label error, and used it to iteratively train a machine learning classifier (Random Forests) using an active learning approach, in which we used the uncertainty in model predictions to select more informative training samples. Minimizing label errors improved model F1 scores by up to 25%, while active learning increased the F1 score by an average of 9.1% between first and last training iterations, which was 2.3% higher than models trained with randomly selected labels. We used the resulting <4 m map of cropland probabilities within a final segmentation step to delineate crop field boundaries. Using an independent map reference sample (n=1,207), we found that overall accuracies of the cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%. Relying on the map reference sample to generate an unbiased area estimate from the field boundary map, we calculate that cropland covers 17.1 (15.4-18.9%). Using the labellers’ digitized field boundaries to estimate biases in the size and number of segments in the field boundary map, we estimate that there are 1,662,281 crop fields in Ghana with an average size of 1.73 ha. Our results demonstrate an adaptable and transferrable approach for mapping the characteristics of croplands on an annual basis and over national extents, with several features that effectively mitigate the errors inherent in remote sensing of smallholder-dominated agriculture.
